---
title: "Residence time and microcystis"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Residence time data

```{r, warning=FALSE}
#Residence time from DSM2 models

library(tidyverse)
library(lubridate)
library(sf)

#read in the residence time data

res1 = read_csv("data/EX_2020_locRT_Jun27.csv") %>%
  pivot_longer(cols = -SimPeriod, names_to = "Location", values_to = "ResTime") %>%
  mutate(Date = as.Date(SimPeriod, format = "%d%b%Y"), Region = str_extract(Location, "[A-Z]+" ))


ggplot(res1, aes(x = Location, y = ResTime)) + geom_boxplot()+facet_wrap(~Region, scales = "free")

#Longer cuttoff

res1.5 = read_csv("data/EX_2020_locRT.csv") %>%
  pivot_longer(cols = -SimPeriod, names_to = "Location", values_to = "ResTime") %>%
  mutate(Date = as.Date(SimPeriod, format = "%d%b%Y"), Region = str_extract(Location, "[A-Z]+" ))


ggplot(res1.5, aes(x = Location, y = ResTime)) + geom_boxplot()+facet_wrap(~Region, scales = "free")


```

OK, so most regions all have fairly similar residence times per reach. But let's look at it by WYT just for funsies

```{r}
Yrs = read_csv("data/yearassignments.csv") %>%
  select(Year, Yr_type, Index)

#In other circumstances I'd merge these by water years, but the october of a year 
#after a dry summer is probalby more like September of the same year than it is like February of the
#next year, so we'll go by calendar year instead.


res1 = mutate(res1, Month = month(Date), Year = year(Date)) %>%
  filter(Region != "SDWS") %>%
  left_join(Yrs)

ggplot(res1, aes(x = Location, y = ResTime)) + geom_boxplot()+
  facet_grid(Yr_type~Region, scales = "free")
```
OK, now let's calculate averages by region

```{r}
resave = group_by(res1, SimPeriod, Region, Date, Yr_type, Month, Year, Index) %>%
  summarize(ResTime = mean(ResTime)) %>%
  mutate(DOY = yday(Date))

ggplot(resave, aes(x = Month, y = ResTime, color = Yr_type)) + geom_point()+
  geom_smooth()+
  facet_wrap(~Region, scales = "free")

```

OK, now let's merge the microcystis data onto the regions. but we need to think about how to do it.
We're trying to use microcystis from the whole time series, but it isn't exactly the same years as the residence time modeling. I think the most logical thing is to average by month and water eyar type.

```{r}
HABregions15 = st_read("data/HABregions15.shp") %>%
  st_transform(crs = 4326) %>%
  st_make_valid()
#summarize residence time by water year type
resave2 = group_by(resave, Region, Yr_type, Month) %>%
  summarize(ResTime = mean(ResTime))  %>%
  rename(RegionDSM = Region)

#switch the names so they  match the other region names
reglook = read_csv("data/regionlookup.csv")

resave2 = left_join(resave2, reglook)

#load HABs data
load("data/HABs.RData")

#Make the data frame a sf object and join it to the regions
HABsReg = st_as_sf(filter(HABs, !is.na(Latitude)), coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_join(HABregions15) %>%
  mutate(Year = case_when(Source == "FMWTx" ~ 2021,
                          TRUE ~ Year)) %>%
  left_join(Yrs) 

#join restime and HABs by region and water year type
HABrestime = left_join(HABsReg, resave2) %>%
  filter(!is.na(Microcystis), Source != "DOP", !is.na(Region), Month %in% c(6,7,8,9,10)) %>%
  mutate(Mic = factor(round(Microcystis), levels = c(1,2,3,4,5), labels = c("Absent", "Low", "Med", "High", "V.High")),
         Yr_type = factor(Yr_type, levels = c("Critical", "Dry", "Below Normal", "Above Normal", "Wet")))

save(HABrestime, file = "data/HABrestime.RData")

#boxplot of residence time by microcystis scores
ggplot(HABrestime, aes(x =Mic, y = ResTime, color = Yr_type)) + 
  geom_boxplot()+
  facet_wrap(~Region)

#scatterplot of microcystis scores by residence time
ggplot(HABrestime, aes(x =ResTime, y = Microcystis, color = Yr_type)) + 
  geom_point(position = "jitter")+
  facet_wrap(~Region, scales = "free_x")


```
OK, not let's try everything together/

```{r}

HABsrestime2017 = filter(HABrestime, Year >2016, !is.na(Mic), !is.na(Region))

#bar plot of frequency of microcystis by water year type and region 
ggplot(HABsrestime2017, aes(x =Yr_type, fill = Mic)) + 
  geom_bar()+
  facet_wrap(~Region)

#percentage version
ggplot(HABsrestime2017, aes(x =Yr_type, fill = Mic)) + 
  geom_bar(position = "fill", color = "black")+
  scale_fill_manual(values = c("grey91", "lemonchiffon","goldenrod2", "chocolate1", "firebrick"), name = "Microcystis\nVisual Index")+
  facet_wrap(~Region)+ theme_bw()


ggplot(HABsrestime2017, aes(x =Month, fill = Mic)) + 
  geom_bar(position = "fill", color = "black")+
  scale_fill_manual(values = c("grey91", "lemonchiffon","goldenrod2", "chocolate1", "firebrick"), name = "Microcystis\nVisual Index")+
  facet_wrap(~Region)+ theme_bw()


```
OK, we see that drier year types have more Microcystis than wetter year types. Also a lot of variation across regions, with the central and south delta having the most

```{r}
#now maybe % microcystis by water year type?


HABRT2 = group_by(HABrestime, Yr_type, Region, Month) %>%
  summarise(Restime = mean(ResTime), n = n(), Mic0 = length(which(Mic == "Absent"))/n, 
            MicLow = length(which(Mic %in% c("Low", "Med")))/n, 
            MicHigh = length(which(Mic %in% c("High", "V.High")))/n)

#this is frequency of absence
ggplot(HABRT2, aes(x = Restime, y = Mic0, color = Month, shape = Yr_type))+
  geom_point()+ ylab("Percentage of observations where Microcytis is absent \n (by month and region)")+
  xlab("Residence time"
  )
```

We can see that most of the months that have 100% absence are in wet years. There aren't as many patterns with month of the year as I was expecting, but it's a bit hard to see. Also, the longest residence time months/regions have a lot of microcystis. 

```{r}
#plot absence/presence/low all together?
HABRT3 = pivot_longer(HABRT2, cols = c(Mic0, MicLow, MicHigh), names_to = "MicLevel", values_to = "Percentage") %>%
  mutate(MicLevel = factor(MicLevel, levels = c("Mic0", "MicLow", "MicHigh")))

ggplot(HABRT3, aes(x = Restime, y = Percentage, color = Month, shape = Yr_type))+
  geom_point()+ ylab("Percentage of observations where Microcytis is absent \n (by month and region)")+
  xlab("Residence time"
  )+
  facet_wrap(~MicLevel)

```

OK, so maybe there is a pattern here? It's honestly really hard to tell. 

```{r}
library(ordinal)
library(effects)
library(MASS)

#Let's try a model of HABs versus temperature, and residence time
HABrestime = mutate(HABrestime, MicF = case_when(Microcystis ==1 ~ "Absent",
                                                 Microcystis %in% 2:3~ "Mid",
                                                 Microcystis %in% 4:5 ~ "High"),
                    MicF = factor(MicF, levels = c("Absent", "Mid", "High"), ordered = T))

#unfortunately, we are missing a lot of temperature  values
HABtest = filter(HABrestime,!is.na(Temperature), !is.na(ResTime), !is.na(Station), !is.na(MicF))


m1 = clmm(MicF ~ ResTime+ Temperature+  (1|Station)+ (1|Year), data = HABtest)

summary(m1)

#not as exciting as I'd think
#Also, i need to work on how to plot this.

HABtest$fits = fitted(m1)

ggplot(HABtest, aes( ResTime, fits, color = MicF))+ geom_point()+ geom_smooth()+
  ylab("Probability")+ xlab("Residence TIme")

#well, that's a mess

```

check

```{r}
ggplot(HABsReg, aes(x = Year))+geom_bar()+
  facet_wrap(~Region)
```



## Dry years only

```{r}
#maybe if we standardized by water year type in some way?
#Let's try just dry and critical years

dry = filter(HABtest, Yr_type %in% c("Dry", "Critical")) 

m1 = clmm(MicF ~ ResTime+ Temperature+ Month+  (1|Station)+ (1|Year), data = dry)

summary(m1)

#not as exciting as I'd think
#Also, i need to work on how to plot this.

dry$fits = fitted(m1)

ggplot(dry, aes( ResTime, fits, color = MicF))+ geom_point()+ geom_smooth(method = "lm")+
  ylab("Probability")+ xlab("Residence TIme")


ggplot(dry, aes(Temperature, fits, color = MicF))+ geom_point()+ geom_smooth(method = "lm")+
  ylab("Probability")+ xlab("Temperature")

ggplot(dry, aes(as.factor(Month), fits, fill = MicF))+ geom_boxplot()+
  ylab("Probability")+ xlab("Month")


#It's definitely better with Month in there

```
```{r}


```



what if we did a binomial model by region?

```{r}

micbin = group_by(dry, Region, Year) %>%
  summarize(Present = length(Microcystis[which(Microcystis>1)]), Absent = length(Microcystis[which(Microcystis==1)]),
            ResTime = mean(ResTime), Temperature = mean(Temperature, na.rm =T), Salinity = mean(Salinity, na.rm = T))

ggplot(micbin, aes(x = ResTime, y = Present/(Present+Absent))) + geom_text(aes(label = Region, color = as.factor(Year)))+
  ylab("Percent of summer dry year observations with Microcystis")+ xlab("residence time (days)")+
  geom_smooth(method = "lm")+ theme_bw()

micbin2 = group_by(dry, Region) %>%
  summarize(Present = length(Microcystis[which(Microcystis>1)]), Absent = length(Microcystis[which(Microcystis==1)]),
            ResTime = mean(ResTime), Temperature = mean(Temperature, na.rm =T), Salinity = mean(Salinity, na.rm = T))

ggplot(micbin2, aes(x = ResTime, y = Present/(Present+Absent))) + geom_text(aes(label = Region))+
  ylab("Percent of summer dry year observations with Microcystis")+ xlab("residence time (days)")+
  geom_smooth(method = "lm")+ theme_bw()

```
I don't hate this.

```{r}
library(lme4)
library(lmerTest)
bn1 = glm(cbind(Present, Absent)~ Temperature + ResTime+ Salinity + Year, data = micbin, family = "binomial")
summary(bn1)
plot(bn1)

micbin = mutate(micbin, Year2 = Year-2000, Temp = scale(Temperature), Res = scale(ResTime), Sal = scale(Salinity))
bn2 = glmer(cbind(Present, Absent)~ Temp + Res+ Salinity + (1|Year2), data = micbin, family = "binomial")
bn2.1 = glmer(cbind(Present, Absent)~ Temp + log(ResTime)+ Salinity + (1|Year2) + (1|Region), data = micbin, family = "binomial")
#bn2 = glm(cbind(Present, Absent)~ Temp + Res+ Salinity + Year2, data = micbin, family = "binomial")
summary(bn2)
plot(bn2)
summary(bn2.1)
plot(bn2.1)
hist(residuals(bn2))
hist(residuals(bn2.1))

#OK, better!

```
```{r}
library(effects)
plot(allEffects(bn2))
```

```{r}
library(DHARMa)
res = simulateResiduals(bn2)
plot(res)
res2 = simulateResiduals(bn2.1)
plot(res2)
#Ok, that is definitely not good

testOutliers(bn2)
testResiduals(res)
testDispersion(res)
plotConventionalResiduals(bn2)
plotConventionalResiduals(bn2.1)
test = residuals(res, quantileFunction = qnorm)
#so we are overdispersed. it's a mess.

plot(density(resid(bn2, type='pearson')))

scatter.smooth(predict(bn2, type='response'), resid(bn2, type='deviance'), col='gray')
#that looks really good though. So i'm confused. 

qqnorm(resid(bn2))
qqline(resid(bn2))
qqnorm(resid(bn2.1))
qqline(resid(bn2.1))

#log-transformed reslidence time makes no difference. Odd. 
#why does this look so much better than the Dharma results? Still not great though. 
```

```{r}
hist(resave2$ResTime)
hist(log(resave2$ResTime))

```




Plot months by region

```{r}
HABsReg = filter(HABsReg, !is.na(Microcystis))
ggplot(HABsReg, aes(x = Month, fill = as.factor(Microcystis)))+ geom_bar(position = "fill")+
  facet_wrap(~Region)
```

```{r}

bn3 = glm(cbind(Present, Absent)~ Temperature + ResTime+ Salinity, data = micbin2, family = "binomial")

#bn2 = glm(cbind(Present, Absent)~ Temp + Res+ Salinity + Year2, data = micbin, family = "binomial")
summary(bn3)
plot(bn3)
hist(residuals(bn3))

#OK, qq plot looks better, other stuff is a mess. Super outliers. 

```

