---
title: "More DSM residence time analyss"
author: "Rosie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(sf)
library(ordinal)
```


My models were all a bit of a mess. It's not really working very well with the residence time values. The salinity relationship makes no sense. I'm going to limit the datast to just the past five years. I may want to limit it to just the delta and kick out SUisun


```{r}
load("HABrestime.RData")

HABres2 = HABrestime %>%
  st_drop_geometry() %>%
  filter(Year>2014, !is.na(Temperature), !is.na(ResTime), !is.na(Region),
                 !is.na(Station), !is.na(Microcystis), Month %in% c(6:10)) %>%
  mutate(MicF = case_when(Microcystis ==1 ~ "Absent",
                                                 Microcystis %in% c(2, 2.5, 3, 3.5)~ "Mid",
                                                 Microcystis %in% c(4, 4.5,5) ~ "High"),
                    MicF = factor(MicF, levels = c("Absent", "Mid", "High"), ordered = T))
  


#boxplot of residence time by microcystis scores
ggplot(HABres2, aes(x =Mic, y = ResTime, color = Yr_type)) + 
  geom_boxplot()+
  facet_wrap(~Region)

#scatterplot of microcystis scores by residence time
ggplot(HABres2, aes(x =ResTime, y = Microcystis, color = Yr_type)) + 
  geom_point(position = "jitter")+
  facet_wrap(~Region, scales = "free_x")


```

Now an ordered regression

```{r}

m1 = clmm(MicF ~ ResTime+ Temperature+ Month + (1|Station)+ (1|Year), data = HABres2)


summary(m1)

#not as exciting as I'd think
#Also, i need to work on how to plot this.

HABres2$fits = fitted(m1)

ggplot(HABres2, aes( ResTime, fits, color = MicF))+ geom_point()+ geom_smooth(method = "lm")+
  ylab("Probability")+ xlab("Residence TIme")


ggplot(HABres2, aes(Temperature, fits, color = MicF))+ geom_point()+ geom_smooth(method = "lm")+
  ylab("Probability")+ xlab("Temperature")

ggplot(HABres2, aes(as.factor(Month), fits, fill = MicF))+ geom_boxplot()+
  ylab("Probability")+ xlab("Month")

ggplot(HABres2, aes(x = Temperature, y = Microcystis))+ geom_point()+ 
  facet_wrap(~Region)

ggplot(HABres2, aes(x = MicF, y = Temperature))+ geom_boxplot()+ 
  facet_wrap(~Region)+ geom_hline(yintercept = 25)


```

UGH. Still not great

Binomial presence absence by region

```{r}
library(lme4)
library(lmerTest)



micbin = group_by(HABres2, Region, Year, Month) %>%
  summarize(Present = length(Microcystis[which(Microcystis>1)]), Absent = length(Microcystis[which(Microcystis==1)]),
            ResTime = mean(ResTime), Temperature = mean(Temperature, na.rm =T), Salinity = mean(Salinity, na.rm = T)) %>%
            mutate(Year2 = Year-2000, Temp = scale(Temperature), Res = scale(ResTime), Sal = scale(Salinity))


bn2 = glmer(cbind(Present, Absent)~ Temp + Res + Month+ Sal+(1|Year2), data = micbin, family = "binomial")
bn2.1 = glmer(cbind(Present, Absent)~ Temp + log(ResTime) + Sal+ (1|Year2) + (1|Region), data = micbin, family = "binomial")
#bn2 = glm(cbind(Present, Absent)~ Temp + Res+ Salinity + Year2, data = micbin, family = "binomial")
summary(bn2)
plot(bn2)
summary(bn2.1)
plot(bn2.1)
hist(residuals(bn2))
hist(residuals(bn2.1))

#Huh. REsiduals look better but other plots look worse with the log0transformation
library(effects)
plot(allEffects(bn2))

summary(micbin)

#Residence time just really isn't a good predictor. 
library(DHARMa)

res1 = simulateResiduals(bn2.1)
plot(res1)

#OK. Log-transformed residnece time is definitely better, if not perfect

```

I don't think the salinity is doin gmuch for us. it's showing an increase with increasing salinity. 

```{r}

library(MuMIn)
micbin = filter(micbin, !is.na(Sal))
global = glmer(cbind(Present, Absent)~ Temp + Res + Month+ Sal+(1|Year2), data = micbin, family = "binomial", na.action = "na.fail")
dredge(global)

#the best model is just temperature and month. 

global = glmer(cbind(Present, Absent)~ Temp + Res + Month+ Sal+ Region +(1|Year2), data = micbin, family = "binomial", na.action = "na.fail")
dredge(global)

#if you add region in there, salinity also pops up, but not super strong

plot(allEffects(global))

ggplot(micbin, aes(x = Temperature, y = Present))+ geom_point()+ geom_smooth()


micbin2017 = filter(micbin, !is.na(Sal), Year > 2016)
global17 = glmer(cbind(Present, Absent)~ Temp + Res+  Month+ Sal+(1|Year2), data = micbin2017, family = "binomial", na.action = "na.fail")
dredge(global17)

best17 = glmer(cbind(Present, Absent)~ Temp + Month+ Res+ (1|Year2), data = micbin2017, family = "binomial", na.action = "na.fail")
summary(best17)

```


```{r}
#try adding a random effect of region

library(MuMIn)
micbin = filter(micbin, !is.na(Sal))
global = glmer(cbind(Present, Absent)~ Temp + Res+(1|Region) + Month+ Sal+(1|Year2), data = micbin, family = "binomial", na.action = "na.fail")
dredge(global)
```

it might be more appropriate to do 2017 onwards, since that's when NCRO started

```{r}
micbin2017 = filter(micbin, !is.na(Sal), Year > 2016)
global17 = glmer(cbind(Present, Absent)~ Temp + Res+(1|Region) + Month+ Sal+(1|Year2), data = micbin2017, family = "binomial", na.action = "na.fail")
dredge(global17)

best17 = glmer(cbind(Present, Absent)~ Temp + Region + Month+ (1|Year2), data = micbin2017, family = "binomial", na.action = "na.fail")
summary(best17)

```

Look at data by region, compare physical variables.

```{r}



physreg = group_by(HABres2, Region, Year) %>%
  summarize(Present = length(Microcystis[which(Microcystis>1)]), Absent = length(Microcystis[which(Microcystis==1)]), Percent = Present/(Present+Absent),
            ResTime = mean(ResTime), Temperature = mean(Temperature, na.rm =T), Salinity = mean(Salinity, na.rm = T),
            Secchi = mean(Secchi, na.rm =T), Depth = mean(Depth, na.rm =T)) %>%
  group_by(Region) %>%
  summarise(Present = mean(Present), Absent = mean(Absent), Percent =Present/(Present+Absent), ResTime = mean(ResTime), Temperature = mean(Temperature, na.rm =T), Salinity = mean(Salinity, na.rm = T), Secchi = mean(Secchi, na.rm =T), Depth = mean(Depth, na.rm =T))

ggplot(physreg, aes(x = Region, y = Salinity))+ geom_col()

ggplot(physreg, aes(x = Region, y = ResTime))+ geom_col()
ggplot(physreg, aes(x = Region, y = Temperature))+ geom_col()
ggplot(physreg, aes(x = Percent, y = Salinity))+ geom_point()+ geom_smooth()
ggplot(physreg, aes(x = Percent, y = Temperature))+ geom_point()+ geom_smooth()
ggplot(physreg, aes(x = Percent, y = ResTime))+ geom_point()+ geom_smooth()

```
Let's try an NMDS plot

```{r}
library(vegan)
physreg = mutate(physreg, Depth = case_when(is.nan(Depth) ~ 8,
                                            TRUE ~ Depth))
regmat = as.matrix(select(physreg, Temperature, Salinity, ResTime, Depth, Secchi)) 

mds1 = metaMDS(regmat, autotransform =T)

plot(mds1, type = "t")

data.scores <- as.data.frame(scores(mds1)$sites)  #Using the scores function from vegan to extract the site scores and convert to a 
data.scores$Region = physreg$Region
data.scores$Percent = physreg$Percent
vars = as.data.frame(scores(mds1)$species)
vars$Label = row.names(vars)
  
ggplot(data.scores, aes(x = NMDS1, y = NMDS2))+ geom_point(aes(size = Percent), alpha = 0.5, color = "lightskyblue")+
  geom_text(aes(label = Region))+ 
  geom_text(data = vars, aes(label = Label), color = "red")+
  geom_segment(data = vars, aes(x = 0, y =0, xend = NMDS1, yend = NMDS2), 
               arrow = arrow(length = unit(0.1,"cm")), inherit.aes = FALSE, 
               color = "red")+
  scale_size_continuous(range = c(1,10), name = "Percent\nof samples \nwith Microcysits\nPresent")+
  theme_bw()
  

```

